Stack Safety is a Security Property

### 2021-04-15 tex/main.tex

\paragraph*{Projections}

We project out the machine state of an MPC-state with \(\pi_\mach\), the
policy state with \(\pi_\pol\), and the context with \(\pi_\context\). We overload
these operators to project out of an machine-trace a trace of machine states,
policy states, and contexts, respectively.

\paragraph*{Head and Last}

We take the first element of a trace with \(\head(\Trace)\), which is a total
function since traces are non-empty, and the final element (if one exists) with
\(\last(\Trace)\), which is partial.

\paragraph*{Until}
The operation \(\PUT ~ f ~ \Trace\) takes a trace
\(\Trace \in \TRACE{\TRACEELEMS %\times \OBSS
}  \) and a predicate
on elements \(f \subset \TRACEELEMS\) and gives the prefix of
\(\Trace\) ending with the first element on which \(f\) holds.

\begin{center}
  \begin{minipage}{.3\textwidth}
    \judgment{\(f ~ \traceelem\)}
             {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\)}
  \end{minipage}
%
  \begin{minipage}{.3\textwidth}
    \judgment{}
             {\(\PUT ~ f ~ \traceelem = \traceelem\)}
%
  \end{minipage}
  \begin{minipage}{.3\textwidth}
    \judgmenttwo{\(\neg f ~ \traceelem\)}{\(\PUT ~ f ~ \Trace = \Trace'\)}
                {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\Trace'\)}
  \end{minipage}
\end{center}
%
\apt{If we are writing \(f \subset \TRACEELEMS\), perhaps we should also
  be writing \(a \in f\), etc.  rather than \(f a\), etc.?}
If the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate always holds, then
\(\PUT\) is the identity function.

% The {\it join} operation, written \(T_1 \cdot T_2\) is the reverse of
%$\mathit{split}$: it takes two traces and links them by replacing the
%last element of the first (if it has one) with the first element of
%the second

%\judgment{} {\(t \cdot T = T\)}
%\judgment{\(T_1 \cdot T_2 = T\)}{\(tT_1 \cdot T_2 = tT\)}
%\leo{Lemma/theorem: split/join = id?}
%\leo{TODO: FORMATTING: Remove force line break/par at end of judgments}

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
If \(f\) is a predicate on machine states, we define \(\mpcstatename
\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
running \(\mpcstatename\) up to \(f\)'') and similarly\apt{??} for traces with policies:

\begin{center}
\judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
            {\(\mpcstatename \hookrightarrow \MPCT | f\)}
\end{center}

@@

In particular, we wish to formulate properties about consecutive sub-traces
in which the context obeys a given predicate. This will commonly be used to extract
individual calls from a trace by taking sub-traces in which the stack is
at or below a given depth\bcp{needed??}. If \(f \subseteq \MACHS \times \CONTEXTS\), then
we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''

\begin{center}
  \judgmenttwo[Now]
              {\(f ~ \mach ~ \context\)}
              {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
              {\(\mach,\pol,\context \hookrightarrow \MPCT \langle f \rangle\)}
%
  \judgmentthree[Skip]
                {\(f ~ \mach ~ \context\)}
                {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
%
  \judgmentthree[Search]
                {\(\neg f ~ \mach ~ \context\)}
                {\(\mach,\pol,\context \hookrightarrow \MPCT | f\)}
                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
\end{center}

@@

We often want to quantify over all traces from any initial state so that our
properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
is {\em reachable} if for some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT\)
and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a sub-trace
such that from some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).
\apt{Unclear how there can be multiple initial states.}

@@

\definition For any depth \(\depth\), define \(f_\depth ~ \mach ~ (\_, \rts)\) to be
\(|\rts| \geq \depth\). Then each reachable segment \(\MPCT \langle f_\depth \rangle\)
corresponds to a trace of a function from its call to its return.
Given such a segment \(S\), let \(\mach,\_,(\domMap,\_)\)  = \(\head(S)\) and let
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
A system enjoys \emph{stepwise stack confidentiality} if, for all \(\MPCT\langle f_\depth\rangle\)
with associated \(m\) and \(\components\) as above, and
for all \(\nach\) such that \(\mach \approx_\components \nach\) and
 \(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\),
 \(\MPCT\langle f_\depth\rangle \doteq \NPCT\langle f_\depth\rangle \).\apt{I found the previous
   version hopelessly confusing, but I'm not at all sure that this one is right! SNA please check.}
%% \definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
%% \(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
%% let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
%%  \(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
%% \(\MPCT\) corresponds to the trace of a function from its call to its return.
%% Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
%% \(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
%% {\em stepwise stack confidentiality} if for all such \(\MPCT\) and \(\NPCT\),
%% \(\MPCT \doteq \NPCT\).

@@

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\). Let
\(\components\) be the set of components \(\component\) such that
\(\domMap ~ \component = \unsealed\) or \(\passed{\depth,\depth'}\) where
\(\depth \leq |\rts| \leq \depth'\) .
Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
{\em stepwise stack confidentiality with passing} if for all such \(\MPCT\) and \(\NPCT\),
\(\MPCT \doteq \NPCT\).

@@

\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
\(\components\) be the set of components \(\component\) such that
\(\domMap ~ \component = \instack{\stackid}{\_}\).
Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
{\em stepwise coroutine confidentiality} in the coroutine setting if for all
such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).

@@

So we extend our model with a notion of observations. We abstract over an
observation type \(\obs \in \OBSS\), containing at minimum the silent observation
\(\tau\). Our machine step function now takes a state and returns a state and an observation,
which is carried over into the policy and context step functions.
%
\[\mach \stepstoobs[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
%

@@

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
Let \(\nach\) be a variant such that \(\mach \approx_\components \nach\) and
\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\).

>>>>

We project out the machine state of an MPC-state with \(\pi_\mach\), the
policy state with \(\pi_\pol\), and the context with \(\pi_\context\).
We take the first element of a trace with \(\head(\MPCT)\), which is a total
function since traces are non-empty, and the final element (if one exists) with
\(\last(\MPCT)\), which is partial.

\paragraph*{Until}
The operation \(\PUT ~ f ~ \MPCT\) takes a trace \(\MPCT\)
and a predicate on states \(f ~ \mpcstate \in \MPCS\) and gives the prefix of
\(\MPCT\) ending with the first element on which \(f\) holds.

\begin{center}
  \begin{minipage}{.3\textwidth}
    \judgment{\(f ~ \mpcstatename\)}
             {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\)}
  \end{minipage}
%
  \begin{minipage}{.3\textwidth}
    \judgment{}
             {\(\PUT ~ f ~ \mpcstatename = \mpcstatename\)}
%
  \end{minipage}
  \begin{minipage}{.3\textwidth}
    \judgmenttwo{\(\neg f ~ \mpcstatename\)}{\(\PUT ~ f ~ \MPCT = \MPCT'\)}
                {\(\PUT ~ f ~ \mpcstatename\MPCT = \mpcstatename\MPCT'\)}
  \end{minipage}
\end{center}
%
If the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate always holds, then
\(\PUT\) is the identity function.

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
If \(f\) is a predicate on machine states, we define \(\mpcstatename
\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
running \(\mpcstatename\) up to \(f\)''):
%
\begin{center}
\judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
            {\(\mpcstatename \hookrightarrow \MPCT | f\)}
\end{center}

@@

individual calls from a trace by taking subtraces in which the stack is
at or below a given depth. If \(f\) is still a predicate on states, then
we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''

\begin{center}
  \judgmenttwo[Now]
              {\(f ~ \mpcstatename\)}
              {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
              {\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\)}
%
  \judgmentthree[Skip]
                {\(f ~ \mpcstatename\)}
                {\(\mpcstatename \hookrightarrow \MPCT | \neg f\)}
                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
%
  \judgmentthree[Search]
                {\(\neg f ~ \mpcstatename\)}
                {\(\mpcstatename \hookrightarrow \MPCT | f\)}
                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
                {\(\mpcstatename \hookrightarrow \MPCT' \langle f \rangle\)}
\end{center}

@@

We often want to quantify over all traces from any initial state so that our
properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
is {\em reachable} if for some initial state \(\mpcstatename = \mach,\pol,\context_0\)
where \(\WF ~ \mach,\pol\), \(\mpcstatename \hookrightarrow \MPCT\)
and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
such that from some similarly well-formed initial state \(\mpcstatename\),
\(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).

@@

\definition For any depth \(\depth\), define \(f_\depth ~ \mach ~ (\_, \rts)\) to be
\(|\rts| \geq \depth\). Then each reachable segment \(\MPCT \langle f_\depth \rangle\)
corresponds to a trace of a function from its call to its return.
Given such a segment \(S\), let \(\mach,\_,(\domMap,\_)\)  = \(\head(S)\) and let
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
A system enjoys \emph{stepwise stack confidentiality} if, for all \(\MPCT\langle f_\depth\rangle\)
with associated \(\mach\) and \(\components\) as above, and
for all \(\nach\) such that \(\mach \approx_\components \nach\) and
 \(\nach \hookrightarrow \NPCT\),
 \(\MPCT\langle f_\depth\rangle \doteq \NPCT | \neg f_\depth \).\apt{I found the previous
   version hopelessly confusing, but I'm not at all sure that this one is right! SNA please check.}
 \sna{It's right, but I need to rethink how the M<f> notation works...}

@@

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\). Let
\(\components\) be the set of components \(\component\) such that
\(\domMap ~ \component = \unsealed\) or \(\passed{\depth,\depth'}\) where
\(\depth \leq |\rts| \leq \depth'\) .
Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
{\em stepwise stack confidentiality with passing} if for all such \(\MPCT\) and \(\NPCT\),
\(\MPCT \doteq \NPCT\).

@@

\definition Let \(\stackid\) be a stack identifier, \(\depth\) be any depth, and
\(f_\depth ~ \mach ~ (\_, \rtm, \_, \stackid)\) hold for \(|\rtm ~ \stackid| \geq \depth\).
For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rtm,\yts,\stackid)\). Let
\(\components\) be the set of components \(\component\) such that
\(\domMap ~ \component = \instack{\stackid}{\_}\).
Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
{\em stepwise coroutine confidentiality} in the coroutine setting if for all
such \(\MPCT\) and \(\NPCT\), \(\MPCT \doteq \NPCT\).

@@

So we extend our model with a notion of observations. We abstract over an
observation type \(\obs \in \OBSS\), containing at minimum the silent observation
\(\tau\). Our machine step function now takes a state and returns a state and an observation,
which is carried over into the policy and context step functions.
%
\[\mach \stepsto[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
%

@@

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
Let \(\nach\) be a variant such that \(\mach \approx_\components \nach\) and
\(\nach \hookrightarrow \NPCT | \neg f_\depth \).
